The Document Object Model
The document object model (DOM) is, as previously mentioned, a way of representing the document
independent of browser type. It allows a developer to access the document via a common set of objects,
properties, methods, and events, and to alter the contents of the web page dynamically using scripts.
You should be aware that some small variations are usually added to the DOM by the browser
vendor. So, to guarantee that you don’t fall afoul of a particular implementation, the W3C has
provided a generic set of objects, properties, and methods that should be available in all browsers, in
the form of the DOM standard.
The DOM Standard
We haven’t talked about the DOM standard so far, and for a particular reason: It’s not the easiest
standard to follow. Supporting a generic set of properties and methods has proved to be a very
complex task, and the DOM standard has been broken down into separate levels and sections
to deal with the different areas. The different levels of the standard are all at differing stages of
completion.

Level 0
Level 0 is a bit of a misnomer, because there wasn’t really a level 0 of the standard. This term in fact
refers to the “old way” of doing things—the methods implemented by the browser vendors before the
DOM standard. Someone mentioning level 0 properties is referring to a more linear notation of accessing
properties and methods. For example, typically you’d reference items on a form with the following code:
document.forms[0].elements[1].value = "button1";
We’re not going to cover such properties and methods in this chapter, because they have been
superseded by newer methods.

Level 1
Level 1 is the first version of the standard. It is split into two sections: One is defined as core
(objects, properties, and methods that can apply to both XML and HTML) and the other as HTMLThe Document Object Model
❘ 235
(HTML‐specific objects, properties, and methods). The first section deals with how to go about
navigating and manipulating the structure of the document. The objects, properties, and methods in
this section are very abstract. The second section deals with HTML only and offers a set of objects
corresponding to all the HTML elements. This chapter mainly deals with the second section—
level 1 of the standard.
In 2000, level 1 was revamped and corrected, though it only made it to a working draft and not to a
full W3C recommendation.

Level 2
Level 2 is complete and many of the properties, methods, and events have been implemented
by today’s browsers. It has sections that add specifications for events and style sheets to the
specifications for core and HTML‐specific properties and events. (It also provides sections on
views and traversal ranges, neither of which is covered in this book; you can find more information
at www.w3.org/TR/2000/PR‐DOM‐Level‐2‐Views‐20000927/ and www.w3.org/TR/2000/
PR‐DOM‐Level‐2‐Traversal‐Range‐20000927/ .)

Level 3
Level 3 achieved recommendation status in 2004. It is intended to resolve a lot of the complications
that still exist in the event model in level 2 of the standard, and adds support for XML features,
such as content models and being able to save the DOM as an XML document.
Level 4
In May 2014, DOM level 4 reached candidate recommendation status. It consolidates DOM level 3
with several independent components. At the time of this writing, no modern browser supports
DOM level 4, although that will change in the future.